---
title: "Homework 1"
author: "Data science for high shcool students, Summer 2018"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Run the code in airnow.R

This is an excercise to run the R code already provided in airnow.R. You would need to create an AirNow account from https://docs.airnowapi.org/account/request/ and then get an API key.

Set the API Key into an *environment variable* (read more here https://en.wikipedia.org/wiki/Environment_variable) using the Sys.setenv function. See help for Sys.setenv by typing *?Sys.setenv* on the R Console in R Studio.

If you are able to run the code successfully you would be able to see the chart generated using the ggplot2 package.

## 2. Basic operations with data frames

See code below and try to write the code as indicated in the comment lines.

```{r}
# create a data frame 
people = data.frame(name = c("person1", "person2", "person3"), age = c(10, 20, 30))

# print the dataframe


# load the purrr package

# use the "map2" function to print the dataframe with each entry written as
# name is person1, age is 10


# print a summary of the people dataframe, use the "summary" function
# lookup help of the "summary" function to see usage example

```




## 3 Doing more with the AirNow API example

Use the code in airnow.R example to get data for the last 1 week and plot it. To get the data for previous days you would need to set the "date=" in the URL to a previous date. This would require a "loop" construct but we do not want to use a for loop, we want to do this with a map. There are multiple ways to do this with the map family of functions but a rather simple way is to use map and make it iterate it over an vector of indices from say 1 to 7.

You would also need to use the "head" function to get the data for the day you are interested in.


