---
title: "Homework 1"
author: "Data science for high shcool students, Summer 2018"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Run the code in airnow.R

This is an excercise to run the R code already provided in airnow.R. You would need to create an AirNow account from https://docs.airnowapi.org/account/request/ and then get an API key.

Set the API Key into an *environment variable* (read more here https://en.wikipedia.org/wiki/Environment_variable) using the Sys.setenv function. See help for Sys.setenv by typing *?Sys.setenv* on the R Console in R Studio.

If you are able to run the code successfully you would be able to see the chart generated using the ggplot2 package.

## 2. Basic operations with data frames

See code below and try to write the code as indicated in the comment lines.

```{r}
# create a data frame 
people = data.frame(name = c("person1", "person2", "person3"), age = c(10, 20, 30))

# print the dataframe


# load the purrr package

# use the "map2" function to print the dataframe with each entry written as
# name is person1, age is 10


# print a summary of the people dataframe, use the "summary" function
# lookup help of the "summary" function to see usage example

```




## 3 Doing more with the AirNow API example

We will discuss this in the class today. We want to write code to get AQI data fro the last one week and plot it.
